Automated Topic Naming

### 2012-01-18 ese/ese.tex
Major rewrite of developer clustering section.

% \subsection{Comparing MaxDB and MySQL}

% \label{sec:comparison}
%\XXX{Not sure there's enough here to do that!}
% We observed that MySQL had more topic subsets than MaxDB. 
% MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500. 
% We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
% In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and
%\emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}. 
%\subsection{Research questions}
%This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
%This analysis addresses our research questions: %raised in Section \ref{sec:analysis}:

\subsection{Do NFR Frequencies Change over Time?}
%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
%\subsubsection{Do label frequencies change over time?}

In both MaxDB and MySQL the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

@@

We wanted to see how similar developers were to each other in terms of
the NFRs that they worked on. 
Within this section we focused on PostgreSQL.
We first conducted pairwise $X^2$ (chi-squared)
tests and Kolmogorov-Smirnov tests for each developer pair. We
found that in $3/10$ of these tests, the developers were significantly
different, i.e., their distributions were not similar. This implies
that there are authors who focus on different NFRs or different
proportion of NFRs.

In Figure \ref{fig:authorcluster} we clustered the top developers from
PostgreSQL using the
Ward method of Hierarchical clustering.  This diagram shows that
\texttt{petere}, \texttt{tgl} and \texttt{momjian} form their own
cluster (they are also the most frequent committers). This is
interesting because it means that the important developers are not
committing code that fits the same NFR profile. In terms of the larger
cluster we can see that \texttt{momjian} and \texttt{tgl} are in the
same cluster, but \texttt{petere} is not. The most frequent
committers do not share the same clusters. This implies that
developers do work on different sets of NFRs and have different
software quality focus.

Figure \ref{fig:authorcluster} clustered the authors into $2$ clusters
and $6$ clusters. We chose $6$ because the difference in distance
between entities and centroids was minimal at $6$ clusters, and the
derivative of distance was the highest in that region as well.

It should be noted that much of PostgreSQL development originates from
the mailinglist and the developers are often pushing integrations. If
we compare the global NFR distribution to each author we find that 25\%
 of the authors have a similarity of $0.47$ or less. This is
depicted in Figure \ref{fig:authorsim}, where we can see that not all
authors are as similar to the global trend as others. Of course, we
found that number of commits correlated with an author's similarity to the
global NFR distribution ($0.59$ Pearson).

Our theory is that the less frequent committers are more
focused and less general, thus their distributions of topics are
different than the main developers who commit code in many different
contexts.  The lead developers either ``wear more hats'' or have more
time to ``wear more hats''.

Given the top 15 PostgreSQL developers we found that proportionally many
developers had a focus on NFR topics. Authors with top proportions per NFR
were:
\texttt{dennis} for usability; 
\texttt{scrappy} and \texttt{meskes} for portability;
\texttt{inoue} and \texttt{neilc} for efficiency;
\texttt{jurka} and \texttt{joe} for reliability;
\texttt{thomas} and \texttt{weick} for functionality;
and \texttt{ishii} and \texttt{scrappy} for maintainability.
Given these results we found that there indeed was a focus
of $1/3$ to $1/10$ of commits dedicated to a particular NFR.

One potential confound for this analysis is that we could be
describing developer style instead of developer focus. Developer style
would be a developers likelihood to use or not use terms found within
our dictionaries and training corpus. Further more a top developer, in
terms of frequency of commits, will have more samples and more likely
take on a more general role. We found in a previous
paper~\cite{largechanges} that often the words used to describe a commit
and the actual author of the commit were often correlated. Thus based
on that observation developer style might be a confounding concern.

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\textwidth]{figures/postgresql-author-cluster}
  \caption{PostgreSQL Author Clusters: Authors clustered by cosine
    distance of their NFR contributions}
\label{fig:authorcluster}
\end{figure}

>>>>

\subsection{Do NFR Frequencies Change over Time?}
In both MaxDB and MySQL the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

@@

We wanted to see whether developers in the PostgreSQL project\footnote{Since the MySQL and MaxDB data had poor records for developer ids, we focused on PostgreSQL.} worked on similar NFRs. For example, one developer might be a usability expert, while
another developer is focused on security. We assume that the NFR labels associated with a commit implicitly reflect work on those NFRs. Being able to report on where effort is being expended is useful for tracking productivity (among other things).
%Within this section we focused on PostgreSQL.
We explored this issue by extracting, for each topic in the PostgreSQL dataset, the developer names responsible for the commits used to generate that topic, and 
the NFRs that are associated with the topic. This produced a map which associated developer name with NFR labels. The six NFR labels (\emph{maintainability, usability, efficiency, functionality, portability, reliability}) and the \emph{None} label describe a 7-dimensional space, into which we can position developers, based on the frequency with which each label occurs in the map.

Using this space, we first conducted pairwise $X^2$ (chi-squared)
tests and Kolmogorov-Smirnov tests for each developer pair. We
found that in $3/10$ of these tests, the two developers were significantly
different (i.e. distant in the 7-d space). This implies
that there are developers who focus on different NFRs or different
proportions of NFRs.

In Figure \ref{fig:authorcluster} we clustered the developers from
PostgreSQL using the
Ward method of Hierarchical clustering on the 7-d space.  
This will group developers according to a distance metric. 
We clustered the authors into both $2$-clusters
and $6$-clusters. We chose $6$-clusters because the difference in distance
between entities and centroids was minimal at $6$-clusters, and the
derivative of distance was the highest in that region as well.

\begin{figure}
  \centering
  \includegraphics[width=.9\textwidth]{figures/postgresql-author-cluster}
  \caption{PostgreSQL Author Clusters: Authors clustered by cosine
    distance of their NFR contributions. The outer rectangles (blue) reflect the 2-clusters; the inner rectangles (purple) reflect 6-clusters.}
\label{fig:authorcluster}
\end{figure}

Figure \ref{fig:authorcluster} shows that when we use $6$-clusters, 
\texttt{petere}, \texttt{tgl} and \texttt{momjian} form their own individual
clusters (i.e., are distant from one another in the 7-d space). 
%NEIL - not sure the relevance of this ... needs more explaining. --> (they are also the most frequent committers). 
This is
interesting because it means that the important developers are not
committing code that fits the same NFR profile. In terms of the 2-clusters,
 we can see that \texttt{momjian} and \texttt{tgl} are in the
same cluster, but \texttt{petere} is not. The most frequent
committers do not share the same clusters, even at the coarsest level. 
This implies that
developers in this sample do work on different sets of NFRs and have different
software quality focuses.

% NEIL --> Explain the reason why this sentence is included: 
%It should be noted that much of PostgreSQL development originates from
%the project mailing list, and the developers are often pushing integrations. 
If we compare the global NFR distribution (that is, the relative global frequency of each NFR label)
to each author we find that 25\%
 of the authors have a similarity of $0.47$ or less. In other words, for these authors (gathered to the lower left of Figure  \ref{fig:authorsim}),
 their NFR distribution does not match well with the global distribution.
We found that number of commits correlated with an author's similarity to the
global NFR distribution ($0.59$ Pearson), i.e., the variables ``number of topics associated with a developer'' and ``number of commits'' are not discriminative.

Our working theory is that less frequent committers are more
specialized, e.g., interested in a specific NFR, while the main developers (i.e., frequent committers) either ``wear more hats'' or have more
time to ``wear more hats''. An interesting extension would be to compare this data with 
the PostgreSQL source code files, to see which developer touched which file.

We say that an author (read developer) is ``proportionately interested'' in an NFR if, for all NFRs with which he or she is associated, a given NFR receives the plurality of his or her commits. This is a measure of relative interest and is independent of number of commits (subject to the caveat about frequent committers, above).
If we look at which of the top 15 developers were proportionately interested in a given NFR, we find the associations identified in Table \ref{tbl:devinterest}. Our data also showed that for these top developers, between $1/10$ and $1/3$ of their commits were labelled with a single NFR.

\begin{table}
\begin{tabular}{l|c}
\toprule
\textbf{NFR} & \textbf{Developer} \\
\midrule
\emph{Usability} & \texttt{dennis, neilc} \\
\emph{Portability} & \texttt{scrappy, meskes} \\
\emph{Efficiency} & \texttt{inoue, neilc} \\
\emph{Reliability} & \texttt{jurka, joe} \\
\emph{Functionality} & \texttt{thomas, weick} \\
\emph{Maintainability} & \texttt{scrappy, ishii}\\
\bottomrule
\end{tabular}
	\caption{Developer interest in NFRs}
	\label{tbl:devinterest}
\end{table}

One potential confound for this analysis is that we could be
describing developer style instead of developer focus. Developer style
would be a developer's likelihood to use or not use terms found within
our dictionaries and training corpus in their commit messages. Furthermore, a top developer, in
terms of number of commits, will have more samples, and thus be more likely
take on a more general role. For example, we found in a previous
project~\cite{largechanges} that there was a good correlation between the words used to describe a commit
and the actual author of the commit. 
%were often correlated. Thus based
%on that observation developer style might be a confounding concern.

@@
